Characteristic polynomials of p-adic matrices

### 2017-02-04 charpoly/charpoly.tex

\noindent
{\bf Numerical experiments.}
We have made numerical experiments in \textsc{SageMath}~\cite{sage}
in order to compare the optimal precision obtained with the methods
explained above with the actual precision obtained by the software.
For doing so, we picked a sample of $1000$ random matrices $M$ in 
$M_9(\Q_2)$ where all the entries at the same \emph{relative
precision}.
We recall that, in \textsc{SageMath}, random elements $x \in \Q_p$ are 
generated as follows. We fix an integer $N$ --- the so-called relative
precision --- and generate elements of $\Q_p$ of the shape
$$x = p^v \cdot \big(a + O\big(p^{N+v_p(a)}\big)\big)$$
where $v$ is a random integer generated according to the distribution:
$$\mathbb P [v = 0] = \frac 1 5 \quad ; \quad
\mathbb P [v = n] = \frac 2 {5\cdot |n| \cdot (|n|+1)} \text{ for }
|n| \geq 1$$
and $a$ is an integer in the range $[0, p^N)$, selected uniformly at random.

@@

\noindent
Lemma 3.4 of \cite{caruso-roe-vaccon:14a} now implies that, if the
$N_{i,j}$'s are large enough and sufficiently well balanced, the optimal
precision on the eigenvalue $\lambda$ is $O(p^{N'})$ with:
$$N' = \min_{1 \leq i, j\leq n} \big(N_{j,i} + \val(C_{i,j}(\lambda)) - 
\val(\chi'_M(\lambda))\big)$$
where $C_{i,j}$ denotes as above the $(i,j)$ entry of $\com(X{-}M)$.
Writing
$\com(X{-}M) = \alpha \cdot P V^t \cdot V Q^t \text{ mod } \chi_M$
as in Proposition~\ref{prop:shortcom}, we find:
\begin{align*}
N' & = \val(\alpha(\lambda)) - \val(\chi'_M(\lambda)) \\
& \hspace{2mm} + \min_{1 \leq i, j\leq n} \big(N_{j,i} + 
\val(P_i V(\lambda)^t) + \val(V(\lambda) Q_j^t)\big)
\end{align*}
where $P_i$ denotes the $i$-th row of $P$ and, similarly, $Q_j$
denotes the $j$-th row of $Q$. Note moreover that $V(\lambda)$ is
the row vector $(1, \lambda, \ldots, \lambda^{n-1})$.
By the discussion of \S \ref{}, the exact value of $M$ can be 
determined for a cost of $\softO(n^\omega)$ operations in $K$ and
$O(n^2)$ operations on integers.

>>>>

\noindent
{\bf Numerical experiments.}
We have made numerical experiments in \textsc{SageMath}~\cite{sage}
in order to compare the optimal precision obtained with the methods
explained above with the actual precision obtained by the software.
For doing so, we picked a sample of $1000$ random matrices $M$ in 
$M_9(\Q_2)$ where all the entries are given at the same \emph{relative
precision}.
We recall that, in \textsc{SageMath}, random elements $x \in \Q_p$ are 
generated as follows. We fix an integer $N$ --- the so-called relative
precision --- and generate elements of $\Q_p$ of the shape
$$x = p^v \cdot \big(a + O\big(p^{N+v_p(a)}\big)\big)$$
where $v$ is a random integer generated according to the distribution:
$$\mathbb P [v = 0] = \frac 1 5 \quad ; \quad
\mathbb P [v = n] = \frac 2 {5\cdot |n| \cdot (|n|+1)} \text{ for }
|n| \geq 1$$
and $a$ is an integer in the range $[0, p^N)$, selected uniformly at random.

@@

\noindent
Lemma 3.4 of \cite{caruso-roe-vaccon:14a} now implies that, if the
$N_{i,j}$'s are large enough and sufficiently well balanced, the optimal
precision on the eigenvalue $\lambda$ is $O(p^{N'})$ with:
$$N' = \min_{1 \leq i, j\leq n} \big(N_{j,i} + \val(C_{i,j}(\lambda)) - 
\val(\chi'_M(\lambda))\big)$$
where $C_{i,j}$ denotes as above the $(i,j)$ entry of $\com(X{-}M)$.
Writing
$\com(X{-}M) = \alpha \cdot P V^t \cdot V Q^t \text{ mod } \chi_M$
as in Proposition~\ref{prop:shortcom}, we find:
\begin{align*}
N' & = \val(\alpha(\lambda)) - \val(\chi'_M(\lambda)) \\
& \hspace{2mm} + \min_{1 \leq i, j\leq n} \big(N_{j,i} + 
\val(P_i V(\lambda)^t) + \val(V(\lambda) Q_j^t)\big)
\end{align*}
where $P_i$ denotes the $i$-th row of $P$ and, similarly, $Q_j$
denotes the $j$-th row of $Q$. Note moreover that $V(\lambda)$ is
the row vector $(1, \lambda, \ldots, \lambda^{n-1})$.
By the discussion of \S \ref{}, the exact value of $N'$ can be 
determined for a cost of $\softO(n^\omega)$ operations in $K$ and
$O(n^2)$ operations on integers.
